---
title: "2. Quality control"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
# remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
library(DoubletFinder)


```


# Set working dir

```{r,echo=F,eval=T,message=F,warning=F,error=F}

here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data 


```{r,echo=F,eval=T,message=FALSE,warning=F}
combined <- qs::qread('combined_raw.qsave')

```

# Number of cells of the original data

```{r,echo=F,eval=T,message=FALSE,warning=F}
table(combined$orig.ident)

```





```{r,echo=F,eval=T,message=FALSE,warning=F}

rb.genes <-
  rownames(combined)[grep("^RP[SL][[:digit:]]", rownames(combined))]
percent.ribo <-
  colSums(combined[rb.genes, ]) / Matrix::colSums(combined) * 100
combined <-
  AddMetaData(combined, percent.ribo, col.name = "percent.ribo")

rb.genes

```

# Mitochondrial  genes 

```{r,echo=F,eval=T,message=FALSE,warning=F}

combined <- PercentageFeatureSet(combined, "^MT-", col.name = "percent.mito")

mito_genes <- rownames(combined)[grep("^MT-", rownames(combined))]
mito_genes
```

# Plot QC

Citing from “Simple Single Cell” workflows (Lun, McCarthy & Marioni, 2017): “High proportions are indicative of poor-quality cells (Islam et al. 2014; Ilicic et al. 2016), possibly because of loss of cytoplasmic RNA from perforated cells. The reasoning is that mitochondria are larger than individual transcript molecules and less likely to escape through tears in the cell membrane.”

```{r,echo=F,eval=T,message=FALSE,warning=F, fig.height=12, fig.width=8}
VlnPlot(
  combined,
  features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"),
  group.by = "orig.ident",
  ncol = 2,
  pt.size = 0.1
)
```

# Filter cells and plot after QC

Similar to [Single cell transcriptomic analysis of cardiac differentiation from human PSCs reveals HOPX-dependent cardiomyocyte maturation](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6220122/).

Since we want to perform QC as permissive of QC thresholding as possible, we can revisit QC if downstream clustering cannot be interpreted.

No mito thres was applied since we have many cells have high mito percentage

These cells are removed:

- Cells with above 50% ribosomal reads

- Cells detected less than 200 genes

- Cells with more than 200,000 total number of molecules

- Cells with genes detected in fewer than 3 cells


```{r,echo=F,eval=T,message=FALSE,warning=F}

combined_qc <-
  subset(combined, subset = percent.ribo < 50 &
           nCount_RNA < 200000 & percent.mito < 25)

```


# Number of cells 

```{r,echo=F,eval=T,message=FALSE,warning=F}
print(paste0("before QC: ", ncol(combined)))
table(combined$orig.ident)

print(paste0("after QC: ", ncol(combined_qc)))
table(combined_qc$orig.ident)
```

# Plot data after QC

```{r,echo=F,eval=T,message=FALSE,warning=F, fig.height=12, fig.width=8}
VlnPlot(
  combined_qc,
  features = c("nFeature_RNA", "nCount_RNA", "percent.mito", "percent.ribo"),
  group.by = "orig.ident",
  ncol = 2,
  pt.size = 0.1
)
```

# Cell-cycle scores

We here perform cell cycle scoring. To score a gene list, the algorithm calculates the difference of mean expression of the given list and the mean expression of reference genes. To build the reference, the function randomly chooses a bunch of genes matching the distribution of the expression of the given list. Cell cycle scoring adds three slots in data, a score for S phase, a score for G2M phase and the predicted cell cycle phase.

```{r,echo=F,eval=T,message=FALSE,warning=F}
combined_qc = NormalizeData(combined_qc)
combined_qc <- CellCycleScoring(object = combined_qc, g2m.features = cc.genes$g2m.genes, 
                              s.features = cc.genes$s.genes)
  
VlnPlot(combined_qc, features = c("S.Score", "G2M.Score"), group.by = "orig.ident", 
        ncol = 2, pt.size = 0.1)

```


```{r,echo=F,eval=F,message=FALSE,warning=F}
# Save Seurat object
# qs::qsave is much faster than save RDS

qs::qsave(combined_qc, "combined.qsave") 
#combined <- qs::qread('combined.qsave')

```

```{r,echo=F,eval=F,message=FALSE,warning=F}

# Alternatively, save RDS

saveRDS(combined_qc, "combined.rds")
```

# Session info

```{r,echo=F,eval=T,message=FALSE,warning=F}
sessionInfo()
```

