---
title: "Cell type annotation"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)
library(fgsea)
library(msigdbr)
library(monocle3)
library(batchelor)
library(Matrix.utils)
#BiocManager::install(c('batchelor', 'Matrix.utils'))
```


# Set working dir

```{r,echo=F,eval=T,message=F,warning=F,error=F}
here::i_am("4_cell_type_annotation.Rmd")
here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data

```{r,echo=F,eval=T,message=F,warning=F,error=F}
source("functions.R")
####

#combined <- readRDS("combined.rds")
combined <- qs::qread('combined.qsave')
####

DefaultAssay(combined) <- "RNA"
custom_color <-
  as.character(palette36.colors(36)[-2])[1:length(levels(Idents(combined)))]
Idents(combined) <- combined$seurat_clusters

provided_marker <- read.csv("provided_marker_manual.csv", header = T)
provided_marker$cell_type <- as.factor(provided_marker$cell_type)

```


# Identify cell clusters

```{r,echo=F,eval=F,message=FALSE,warning=F}

dbs <-
  c(
    "GO_Molecular_Function_2018",
    "GO_Cellular_Component_2018",
    "GO_Biological_Process_2018",
    "KEGG_2019_Human"
  )

DefaultAssay(combined) <- "RNA"
#combined <- FindClusters(combined, resolution = 0.5)

```


```{r,echo=F,eval=T,message=T,warning=F,error=F}
table(combined$seurat_clusters)

#Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")

DimPlot(combined, reduction = "umap", cols = as.vector(palette36.colors(36)[-2]),label = T)

#DimPlot(combined, reduction = "tsne", cols = as.vector(palette36.colors(36)[-2]),label = T)

```


```{r,echo=F,eval=F,message=F,warning=F,error=F}

p1 <-
  Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")
png(
  paste("Predicted_clusters_umap.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
p1
dev.off()
p1


```


# Cluster-specific genes

The Seurat package FindAllMarkers function was used to identify cluster-specific genes. For example, for cluster 0, the package performs differential testing using cells from cluster 0 vs all other cells. Only genes with log2 fold-change > 0.5 (logfc.threshold=0.5), and genes that are detected in a minimum fraction of 20% (min.pct=0.2) cells in either of the two populations were saved.

Considering too many tables or figures would drastically slow down the rendering of this report. I saved the Cluster-specific genes related results in the 'cluster_specific_genes' folder.

```{r,echo=F,eval=F,message=F,warning=F,error=F}
library(future)
plan("multiprocess", workers = 20)
plan()
getwd()
Idents(combined) <- combined$seurat_clusters
DefaultAssay(combined) <- "RNA"
# 30 mins
#cts_markers <- FindAllMarkers(combined, logfc.threshold = 0.5, min.pct = 0.2)
cts_markers %>%
  filter(p_val_adj < 0.05) %>%
  write_csv("cluster_specific_genes_combined.csv")
i = 1
for (i in seq_along(levels(cts_markers$cluster))) {
  this_cluster <- levels(cts_markers$cluster)[i]
  cts_markers %>%
    filter(cluster == this_cluster & p_val_adj < 0.05) %>%
    write_csv(paste0("marker_cluster_", this_cluster, ".csv"))
}


```

# Cluster enrichment analysis

Enrichment results for each cluster were saved at 'cluster_enrichment' folder.

```{r,echo=F,eval=F,message=F,warning=F,error=F, fig.width=12,fig.height=8}
#cts_markers <- read.csv("cluster_specific_genes_combined.csv")

i=1
DefaultAssay(combined) <- "RNA"
Idents(combined) <- combined$seurat_clusters

for (i in seq_along(levels(Idents(combined)))) {
  Idents(combined) <- combined$seurat_clusters
  this_subset <- subset(combined, idents = levels(Idents(combined))[i])
  DefaultAssay(this_subset) <- 'RNA'
  
  res <- cts_markers %>%
    dplyr::filter(cluster == levels(Idents(combined))[i]) %>%
    dplyr::select(gene, avg_log2FC) %>%
    arrange(desc(avg_log2FC)) %>%
    na.omit() %>%
    distinct() %>%
    group_by(gene) %>%
    deframe() %>%
    sort(decreasing = T)
    
 
  m_df = msigdbr(species = "Homo sapiens", category = "H")
  m_list = m_df %>% split(x = .$gene_symbol, f = .$gs_name)
  
  fgseaRes <- fgsea(pathways=m_list, stats=res, nperm=1000)
  fgseaResTidy <- fgseaRes %>%
    as_tibble() %>%
    arrange(desc(NES)) %>% 
    dplyr::select(-leadingEdge, -ES, -nMoreExtreme) %>% 
    arrange(padj) %>%
    write_csv(paste0("GSEA_cluster_",levels(Idents(combined))[i],".csv")) 
  
}
```


# Clusters UMAP plot split by condition

## Split by condition 

Number of cells for each cluster:
 
```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}
condition_idents <- as.factor(combined$orig.ident)
levels(condition_idents) <- c(rep("ctrl",6), rep("N1KO",6))
combined <- AddMetaData(combined, condition_idents, col.name = "condition")

table(combined$condition, combined$seurat_clusters)

```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

Idents(combined) <- combined$seurat_clusters

p2 <-
  DimPlot(
    combined,
    reduction = "umap",
    split.by = "condition",
    cols = as.vector(palette36.colors(36)[-2]),
    ncol = 2,
    pt.size = 0.7
  )

png(
  paste("cluster_split_by_condition.png", sep = ""),
  width = 3000,
  height = 1500,
  res = 300
)

p2
dev.off()
p2

```

## Split by samples

Number of cells for each cluster:


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

table(combined$orig.ident, combined$seurat_clusters)

```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=8,fig.height=16}



p3 <-
  DimPlot(
    combined,
    reduction = "umap",
    group.by = "seurat_clusters",
    split.by = "orig.ident",
    cols = as.vector(palette36.colors(36)[-2]),
    ncol = 6,
    pt.size = 0.6
  )

png(
  paste("cluster_split_by_sample.png", sep = ""),
  width = 6000,
  height = 2000,
  res = 300
)

p3
dev.off()
p3

```


# Annotate by cell type markers

```{r,echo=F,eval=F,message=FALSE,warning=F}

Idents(combined) <- combined$orig.ident
DefaultAssay(combined) <- "RNA"

# seq_along(levels(provided_marker$cell_type))
for (idx in seq_along(levels(provided_marker$cell_type))) {
  this_ct <- levels(provided_marker$cell_type)[idx]
  
  this_markers <- provided_marker %>%
    filter(cell_type == this_ct) %>%
    pull(gene)
  this_direction <- provided_marker %>%
    filter(cell_type == this_ct) %>%
    pull(direction)
  this_data <- Seurat::FetchData(object = combined, vars = this_markers)
  selected_cells <- rownames(this_data)
  for (k in seq_along(this_markers)) {
    if (this_direction[k] == "pos") {
      this_cells <- this_data %>%
        filter(get(this_markers[k]) >  0.1) %>%
        rownames_to_column("cell") %>%
        pull(cell)
      selected_cells <- intersect(selected_cells, this_cells)
    } else{
      this_cells <- this_data %>%
        filter(get(this_markers[k]) <= 0.1) %>%
        rownames_to_column("cell") %>%
        pull(cell)
      selected_cells <- intersect(selected_cells, this_cells)
    }
  }
  
  selected_cells <- unique(selected_cells)
  this_obj <- combined[, selected_cells]
  
  combined@meta.data$tmp_ct <-
    ifelse(rownames(combined@meta.data) %in% colnames(this_obj),
           "positive",
           "negative")
  
  
  #custom_color <- c('#C0C0C0', '#B00D23')
  Idents(combined) <- combined$orig.ident
  this_ct_ident <- gsub(" ", "_", this_ct)
  tmp_ct_index <- which(colnames(combined@meta.data) == 'tmp_ct')
  colnames(combined@meta.data)[tmp_ct_index] <- this_ct_ident

  
}




```


```{r,echo=F,eval=F,message=F,warning=F,error=F}
qs::qsave(combined,'combined.qsave')
```


```{r,echo=F,eval=T,message=F,warning=F,error=F}

```
