---
title: "Cell type annotation"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(dplyr)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)
library(fgsea)
library(msigdbr)
```


# Set working dir

```{r,echo=F,eval=T,message=F,warning=F,error=F}

here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data

```{r,echo=F,eval=T,message=F,warning=F,error=F}
source("functions.R")
####

#combined <- readRDS("combined.rds")
combined <- qs::qread('combined.qsave')
####

DefaultAssay(combined) <- "RNA"
custom_color <-
  as.character(palette36.colors(36)[-2])[1:length(levels(Idents(combined)))]
Idents(combined) <- combined$seurat_clusters
```

# Identify cell clusters

```{r,echo=F,eval=F,message=FALSE,warning=F}

dbs <-
  c(
    "GO_Molecular_Function_2018",
    "GO_Cellular_Component_2018",
    "GO_Biological_Process_2018",
    "KEGG_2019_Human"
  )

DefaultAssay(combined) <- "RNA"
#combined <- FindClusters(combined, resolution = 0.5)

```


```{r,echo=F,eval=T,message=T,warning=F,error=F}
table(combined$seurat_clusters)

#Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")

DimPlot(combined, reduction = "umap", cols = as.vector(palette36.colors(36)[-2]),label = T)

#DimPlot(combined, reduction = "tsne", cols = as.vector(palette36.colors(36)[-2]),label = T)

```


```{r,echo=F,eval=F,message=F,warning=F,error=F}

p1 <-
  Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")
png(
  paste("Predicted_clusters_umap.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
p1
dev.off()
p1


```

3D version UMAP may be more clear

```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=10, fig.height=8}
######### 3D UMAP
DefaultAssay(combined) <- "RNA"
# Interacive multimodal 3D UMAP plotting of scRNA sequencing datasets
# The following is a length of code generated to create nice 
# 3D UMAP plots of seurat v3.0.0-v3.1.1 objects utilizing the visualization 
# package plot_ly

# R v3.5.3 (x64 bit) and RStudio v1.2.1335 (x64 bit) were used for running this code :)

# Seurat is a multimodal single Cell RNA seq analysis algorithm created by
# The Satija Lab. For more information please see: https://satijalab.org/seurat/

# Contributors (by their Github handles):
# @Dragonmasterx87 (Dept. of Cell Biology, UM)
# @msaadsadiq (Dept. of Electrical and Computer Engineering, UM)

# Install plot_ly
#install.packages('plotly')

# Load plot_ly
library(plotly)

# Construct a dataframe using data from your pre-clustered Seurat v3.1.1 object
# Here 'seurat_clusters' is list of numeric cluster identities, you can find it here: yourseuratobject[["seurat_cluster"]], 
# or yourseuratobject$seurat_clusters, where 'yourseuratobject' is a Seurat object created with Seurat v3.1.1 (works for v3.0.0 as well)
yourseuratobject <- combined

# Re-run UMAPs that you have accurate calculations for all UMAP(s)
#yourseuratobject <- RunUMAP(yourseuratobject,
#                            dims = 1:30,
#                            n.components = 3L)

# Extract tSNE information from Seurat Object
umap_1 <- yourseuratobject[["umap"]]@cell.embeddings[,1]
umap_2 <- yourseuratobject[["umap"]]@cell.embeddings[,2]
umap_3 <- yourseuratobject[["umap"]]@cell.embeddings[,3]

# Visualize what headings are called so that you can extract them to form a dataframe
# Embeddings(object = yourseuratobject, reduction = "umap")

# Prepare a dataframe for cell plotting
plot.data <- FetchData(object = yourseuratobject, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "seurat_clusters"))

# Make a column of row name identities (these will be your cell/barcode names)
plot.data$label <- paste(rownames(plot.data))

# Plot your data, in this example my Seurat object had 21 clusters (0-20)
fig <- plot_ly(data = plot.data, 
               x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
               color = ~seurat_clusters, 
               colors = custom_color,
               type = "scatter3d", 
               mode = "markers", 
               marker = list(size = 1.5, width=1), # controls size of points
               text=~label, #This is that extra column we made earlier for which we will use for cell ID
               hoverinfo="text") #When you visualize your plotly object, hovering your mouse pointer over a point shows cell names


# Updates stemming from Issue #9 Having a fixed scale on axes while selecting particular clusters
# @rtoddler thanks for the suggestions!
# Before you plot, set the ranges of the axis you desire. This set axis range will be 
# present across all clusters, and plotly will not adjust for axis length anymore
# this axis length will persist even when selecting some clusters

# xaxis
axx <- list(
  nticks = 4,
  range = c(-10,10) #select range of xaxis
)

# yaxis
axy <- list(
  nticks = 4,
  range = c(-10,10) #select range of yaxis
)

#zaxis
axz <- list(
  nticks = 4,
  range = c(-10,10) #select range of zaxis
)

#fig <- fig %>% layout(scene = list(xaxis=axx,yaxis=axy,zaxis=axz))
fig_cube <- fig %>% layout(scene = list(xaxis=axx,yaxis=axy,zaxis=axz, aspectmode='cube')) # To maintain cubic aspect
#fig
fig_cube
```

# NOTCH1 expression


```{r,echo=F,eval=T,message=F,warning=F,error=F}

FeaturePlot(combined, "NOTCH1", pt.size = 0.7)

VlnPlot(combined, "NOTCH1", group.by = "orig.ident")
```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=10, fig.height=8}

# Alternative method as designed by @vertesy (Thanks for the suggestions!)
# create a dataframe
goi <- "NOTCH1"
plotting.data <- FetchData(object = yourseuratobject, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .95)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called ACTB
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c('darkgrey', 'red'), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 1.4), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)

```


# Cluster-specific genes

The Seurat package FindAllMarkers function was used to identify cluster-specific genes. For example, for cluster 0, the package performs differential testing using cells from cluster 0 vs all other cells. Only genes with log2 fold-change > 0.5 (logfc.threshold=0.5), and genes that are detected in a minimum fraction of 20% (min.pct=0.2) cells in either of the two populations were saved.

Considering too many tables or figures would drastically slow down the rendering of this report. I saved the Cluster-specific genes related results in the 'cluster_specific_genes' folder.

```{r,echo=F,eval=F,message=F,warning=F,error=F}
library(future)
plan("multiprocess", workers = 20)
plan()
getwd()
Idents(combined) <- combined$seurat_clusters
DefaultAssay(combined) <- "RNA"
# 30 mins
#cts_markers <- FindAllMarkers(combined, logfc.threshold = 0.5, min.pct = 0.2)
cts_markers %>%
  filter(p_val_adj < 0.05) %>%
  write_csv("cluster_specific_genes_combined.csv")
i = 1
for (i in seq_along(levels(cts_markers$cluster))) {
  this_cluster <- levels(cts_markers$cluster)[i]
  cts_markers %>%
    filter(cluster == this_cluster & p_val_adj < 0.05) %>%
    write_csv(paste0("marker_cluster_", this_cluster, ".csv"))
}


```

# Cluster enrichment analysis

Enrichment results for each cluster were saved at 'cluster_enrichment' folder.

```{r,echo=F,eval=F,message=F,warning=F,error=F, fig.width=12,fig.height=8}
#cts_markers <- read.csv("cluster_specific_genes_combined.csv")

i=1
DefaultAssay(combined) <- "RNA"
Idents(combined) <- combined$seurat_clusters

for (i in seq_along(levels(Idents(combined)))) {
  Idents(combined) <- combined$seurat_clusters
  this_subset <- subset(combined, idents = levels(Idents(combined))[i])
  DefaultAssay(this_subset) <- 'RNA'
  
  res <- cts_markers %>%
    dplyr::filter(cluster == levels(Idents(combined))[i]) %>%
    dplyr::select(gene, avg_log2FC) %>%
    arrange(desc(avg_log2FC)) %>%
    na.omit() %>%
    distinct() %>%
    group_by(gene) %>%
    deframe() %>%
    sort(decreasing = T)
    
 
  m_df = msigdbr(species = "Homo sapiens", category = "H")
  m_list = m_df %>% split(x = .$gene_symbol, f = .$gs_name)
  
  fgseaRes <- fgsea(pathways=m_list, stats=res, nperm=1000)
  fgseaResTidy <- fgseaRes %>%
    as_tibble() %>%
    arrange(desc(NES)) %>% 
    dplyr::select(-leadingEdge, -ES, -nMoreExtreme) %>% 
    arrange(padj) %>%
    write_csv(paste0("GSEA_cluster_",levels(Idents(combined))[i],".csv")) 
  
}
```


# Clusters UMAP plot split by condition

## Split by condition 

Number of cells for each cluster:
 
```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}
condition_idents <- as.factor(combined$orig.ident)
levels(condition_idents) <- c(rep("ctrl",6), rep("N1KO",6))
combined <- AddMetaData(combined, condition_idents, col.name = "condition")

table(combined$condition, combined$seurat_clusters)

```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

Idents(combined) <- combined$seurat_clusters

p2 <-
  DimPlot(
    combined,
    reduction = "umap",
    split.by = "condition",
    cols = as.vector(palette36.colors(36)[-2]),
    ncol = 2,
    pt.size = 0.7
  )

png(
  paste("cluster_split_by_condition.png", sep = ""),
  width = 3000,
  height = 1500,
  res = 300
)

p2
dev.off()
p2

```

## Split by samples

Number of cells for each cluster:


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

table(combined$orig.ident, combined$seurat_clusters)

```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=8,fig.height=16}

Idents(combined) <- combined$seurat_clusters

p3 <-
  DimPlot(
    combined,
    reduction = "umap",
    split.by = "orig.ident",
    cols = as.vector(palette36.colors(36)[-2]),
    ncol = 2,
    pt.size = 0.7
  )

png(
  paste("cluster_split_by_sample.png", sep = ""),
  width = 2000,
  height = 4000,
  res = 300
)

p3
dev.off()
p3

```


# Cell type markers

- The input marker gene list is stored at 'provided_marker.csv'
- The gene expression plots of the cell types are stored in 'marker_plots' folder. 

Some of genes are not presented in the data, I have renamed some of these genes:

- T -> ?

- FLK1 -> KDR

- CD31 (PECAM-1) -> PECAM1

- SMA -> ?

- TE-7 -> ?


```{r,echo=F,eval=F,message=F,warning=F,error=F}
library(SCINA)
library(preprocessCore)
signatures <- preprocess.signatures("provided_marker.csv")

DefaultAssay(combined) <- "RNA"
this_markers <- unlist(signatures)
for (i in seq_along(this_markers)) {
  p_scina <- FeaturePlot(combined, as.character(this_markers[i]), ncol = 1)
  png(
    paste(
      names(this_markers[i]),
      "_marker_",
      as.character(this_markers[i]),
      ".png",
      sep = ""
    ),
    width = 1800,
    height = 1500,
    res = 300
  )
  print(p_scina)
  dev.off()
}


for (i in seq_along(this_markers)) {
  p_scina <- VlnPlot(combined, as.character(this_markers[i]), ncol = 1, group.by = "seurat_clusters")
  png(
    paste(
      names(this_markers[i]),
      "_violin_",
      as.character(this_markers[i]),
      ".png",
      sep = ""
    ),
    width = 1800,
    height = 1500,
    res = 300
  )
  print(p_scina)
  dev.off()
}



for (i in seq_along(this_markers)) {
  p_scina <- VlnPlot(combined, as.character(this_markers[i]), ncol = 1, group.by = "orig.ident")
  png(
    paste(
      names(this_markers[i]),
      "_violin_",
      as.character(this_markers[i]),
      ".png",
      sep = ""
    ),
    width = 1800,
    height = 1500,
    res = 300
  )
  print(p_scina)
  dev.off()
}


```


```{r,echo=F,eval=T,message=F,warning=F,error=F}

```



```{r,echo=F,eval=F,message=FALSE,warning=F}
# Save Seurat object
# qs::qsave is much faster than save RDS

qs::qsave(combined, "combined.qsave") 
#combined <- qs::qread('combined.qsave')

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}
# Manually save the rds, do not save on knitr
saveRDS(combined, "combined.rds")
```

