---
title: "Mintao NOTCH1 scRNA-seq analysis"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(RColorBrewer)
library(Polychrome)
library(ggplot2)
suppressPackageStartupMessages(library(SCINA))
suppressPackageStartupMessages(library(preprocessCore))
library(stringr)
library(tidyverse)
library(enrichR)
library(DT)
library(cowplot)
library(clusterProfiler)
library(scCATCH)
library(org.Mm.eg.db)
library(clusterProfiler)
library(dittoSeq)
library(qs)
#library(ggrepel)
#library(ggfortify)
#library(QUBIC)
#library(DESeq2)
#library(pheatmap)
#library(patchwork)
#library(monocle)
#library(DelayedArray)
#library(DelayedMatrixStats)
#library(org.Mm.eg.db)
#library(VennDiagram)

#
knitr::opts_knit$set(root.dir = "D:/Project/mingtao")

setwd("D:/Project/bai/il27")
```



```{r,echo=F,eval=T,message=F,warning=F,error=F}
#Load data
####

#combined <- readRDS("combined.rds")
combined <- qs::qread('combined.qsave')
####

source("functions.R")
dbs <-
  c(
    "GO_Molecular_Function_2018",
    "GO_Cellular_Component_2018",
    "GO_Biological_Process_2018",
    "KEGG_2019_Mouse"
  )


```

# Background


# Data overview

After preprocessing and data integration, the dataset has 16400 genes across X cells. X control cells and X treated cells.

```{r,echo=F,eval=T,message=F,warning=F,error=F}
table(combined$condition)
tb_all <- table(combined$condition)
#dim(combined)
```



```{r,echo=F,eval=T,message=FALSE,warning=F}

#DefaultAssay(combined) <- "integrated"
#combined <- FindClusters(combined, resolution = 0.2)
DefaultAssay(combined) <- "RNA"
#custom_color <- as.character(palette36.colors(36)[-2])[1:length(levels(Idents(combined)))]
#custom_color <- as.character(palette36.colors(36)[-2])

Idents(combined) <- combined$seurat_clusters
```

# Quality control

Shown in the count table and UMAP plot, 24 cell clusters were predicted by Seurat package. 

```{r,echo=F,eval=T,message=T,warning=F,error=F}
table(combined$seurat_clusters)

Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")

DimPlot(combined, reduction = "umap", cols = as.vector(palette36.colors(36)[-2]),label = T)

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

p1 <-
  Plot.cluster2D(combined, pt_size = 0.4, txt = "Predicted cluster")
png(
  paste("Predicted_clusters.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
p1
dev.off()
p1
```

```{r,echo=F,eval=F,message=F,warning=F,error=F}
library(future)
plan("multiprocess", workers = 20)
plan()
getwd()
Idents(combined) <- combined$seurat_clusters
DefaultAssay(combined) <- "RNA"
# 30 mins
#cts_markers <- FindAllMarkers(combined, logfc.threshold = 0.5, min.pct = 0.2)
cts_markers %>%
  filter(p_val_adj < 0.05) %>%
  write_csv("cluster_specific_genes_combined.csv")
i = 1
for (i in 1:length(levels(cts_markers$cluster))) {
  this_cluster <- levels(cts_markers$cluster)[i]
  cts_markers %>%
    filter(cluster == this_cluster & p_val_adj < 0.05) %>%
    write_csv(paste0("marker_cluster_", this_cluster, ".csv"))
}


```

## Cluster-specific genes & functional gene set enrichment analysis

I used Seurat package FindAllMarkers function to identify cluster-specific genes. For example, for cluster 0, the package performs differential testing using cells from cluster 0 vs all other cells. Only genes with log2 fold-change > 0.5 (logfc.threshold=0.5), and genes that are detected in a minimum fraction of 20% (min.pct=0.2) cells in either of the two populations were saved.

For cluster-specific genes from each cluster (adj p-value < 0.05), the genes were send to Enrichr server to perform functional gene set enrichment analysis. The selected databases were: "GO_Molecular_Function_2018", "GO_Cellular_Component_2018", "GO_Biological_Process_2018", "KEGG_2019_Mouse".

Considering too many tables or figures would drastically slow down the rendering of this report. I saved the Cluster-specific genes related results in the folders:

- cluster_specific_genes_table: Cluster specific genes table saved for each cluster. "cluster_specific_genes_combined.csv" is the combined results of all 17 clusters.
- cluster_specific_genes_plot_top50: Feature plot of top 50 genes from each cluster
- cluster_specific_genes_enrichment: Enrichment analysis table for each clusters



```{r,echo=F,eval=F,message=F,warning=F,error=F}
i = 1
for (i in 1:length(as.character(unique(cts_markers$cluster)))) {
  this_ct_name <- as.character(unique(cts_markers$cluster))[i]
  this_cts_genes <- cts_markers %>%
    filter(cluster == this_ct_name) %>%
    pull(gene)
  
  this_enriched <- enrichr(this_cts_genes, dbs)
  
  write.csv(
    this_enriched$GO_Molecular_Function_2018[, c(-5, -6)],
    paste0("Cluster_", this_ct_name, "_GO_MF.csv")
  )
  write.csv(
    this_enriched$GO_Cellular_Component_2018[, c(-5, -6)],
    paste0("Cluster_", this_ct_name, "_GO_CC.csv")
  )
  write.csv(
    this_enriched$GO_Biological_Process_2018[, c(-5, -6)],
    paste0("Cluster_", this_ct_name, "_GO_BP.csv")
  )
  write.csv(
    this_enriched$KEGG_2019_Mouse[, c(-5, -6)],
    paste0("Cluster_", this_ct_name, "_KEGG.csv")
  )
  
}

```

## DEG between the same clusters in AAV-IL-27 vs AAV-ctrl

A positive log-foldchange mean the gene expression is higher in the IL37 group. The result tables are stored in the '/result_03122021/DEG between the same clusters in AAV-IL-27 vs AAV-ctrl' folder

```{r,echo=F,eval=F,message=F,warning=F,error=F}


Idents(combined) <- combined$seurat_clusters
i = j = 1
for (i in 1:length(levels(Idents(combined)))) {
  this_cluster <- as.numeric(levels(Idents(combined))[i])
  out_filename <-
    paste0("Cluster_", this_cluster, "_IL27_vs_ctrl.csv")
  
  this_markers <-
    FindMarkers(
      combined,
      subset.ident = this_cluster,
      ident.1 = "IL27",
      ident.2 = "Control",
      group.by = "Condition",
      verbose = FALSE
    ) %>%
    rownames_to_column("gene") %>%
    write_csv(out_filename)
}



```

# Clusters UMAP plot split by condition

## Split by condition 

Number of cells for each cluster:
 
```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}
table(combined$Condition, combined$seurat_clusters)

```


```{r,echo=F,eval=F,message=F,warning=F,error=F}
# save top genes feature plots by clusters
TOP = 50

Idents(combined) <- combined$seurat_clusters
i = j = 1
for (i in 1:length(levels(Idents(combined)))) {
  #this_markers <- strsplit(clu_ann$celltype_related_marker[i], ", ")[[1]]
  this_cluster <- levels(cts_markers$cluster)[i]
  dir.create(paste0("Cluster_", this_cluster), showWarnings = F)
  this_markers <- cts_markers %>%
    filter(cluster == this_cluster) %>%
    top_n(TOP) %>%
    pull(gene)
  for (j in 1:length(this_markers)) {
    p_cluster_marker <- FeaturePlot(combined, this_markers[j], ncol = 1)
    png(
      paste(
        "Cluster_",
        this_cluster,
        "_rank_",
        j,
        "_",
        this_markers[j],
        ".png",
        sep = ""
      ),
      width = 1500,
      height = 1500,
      res = 300
    )
    print(p_cluster_marker)
    dev.off()
  }
}

# we have cluster 0 to length-1
for (i in 0:(length(levels(Idents(combined)))-1)) {
  print(i)
  system(paste0("mv Cluster_", i, "_* Cluster_", i))
}

```

```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

Idents(combined) <- combined$seurat_clusters

p2 <-
  DimPlot(
    combined,
    reduction = "umap",
    split.by = "Condition",
    cols = as.vector(palette36.colors(36)[-2]),
    ncol = 2
  )

png(
  paste("cluster_split_by_condition.png", sep = ""),
  width = 4000,
  height = 2000,
  res = 300
)

p2
dev.off()
p2

```

## Split by condition with label


```{r,echo=F,eval=T,message=F,warning=F,error=F}

p2 <- DimPlot(
  combined,
  reduction = "umap",
  split.by = "Condition",
  cols = as.vector(palette36.colors(36)[-2]),
  label = T,
  label.size = 3,
  repel = T
)

p2 
png(
  paste("cluster_split_by_condition_with_label.png", sep = ""),
  width = 4000,
  height = 2000,
  res = 300
)
print(p2)
dev.off()

```

## Cluster split by condition with label box

```{r,echo=F,eval=T,message=F,warning=F,error=F}

p3 <- DimPlot(
  combined,
  reduction = "umap",
  split.by = "Condition",
  cols = as.vector(palette36.colors(36)[-2]),
  label = T,
  label.size = 3,
  label.box = T,
  repel = T
)

p3 
png(
  paste("cluster_split_by_condition_with_label_box.png", sep = ""),
  width = 4000,
  height = 2000,
  res = 300
)
print(p3)
dev.off()
```

# Predict cell type annotation 

I used SCINA and public marker gene database to annotate cell type automatically. This should help annotate cell clusters to cell types. 

This tool is a semi-automatic cell type annotation tool. It uses a marker gene list and gene expression values as input. It achieves good performance from this paper: A comparison of automatic cell identification methods for single-cell RNA sequencing data. The marker genes were collected from CellMatch and PanlaoDB databases. I selected marker genes from spleen & immune systems.

- The input marker gene list has 38 cell types and is stored at 'bai_il27_marker_SCINA.csv'
- The SCINA tool annotated 14 cell types. The gene expression plots of the 14 cell types are stored in 'SCINA' folder. 

In summary, the computational approch may not be reliable, please **double-check the cell type annotations** before we moved to the downstream analysis.


```{r,echo=F,eval=F,message=F,warning=F,error=F}

#combined <- FindClusters(combined, resolution = 0.1)
library(SCINA)
library(preprocessCore)
exp <- as.matrix(GetAssayData(object = combined[["RNA"]], slot = "data"))
# exp <- log1p(exp)
exp[] <- normalize.quantiles(exp)
rownames(exp) <- toupper(rownames(exp))

signatures <- preprocess.signatures("bai_il27_marker_SCINA.csv")
keep_scina_cell_type <- as.numeric(which(!sapply(signatures, function(x) {
  all(!x %in% rownames(exp))
})))
signatures <- signatures[keep_scina_cell_type]

results <- SCINA(exp, signatures,
  max_iter = 100, convergence_n = 20,
  convergence_rate = 0.99, sensitivity_cutoff = 0.9, rm_overlap = F, allow_unknown = T, log_file = "SCINA.log"
)
combined <- AddMetaData(combined, as.factor(results$cell_labels), col.name = "scina_cell_type")

Idents(combined) <- combined$seurat_clusters
annotate_cell_type <- as.factor(combined$seurat_clusters)
annotate_cell_type_name <- vector()
# iterate i as all seurat cell clusters index
for (i in 1:length(levels(Idents(combined)))) {
  this_cluster_cells <- WhichCells(combined, idents = levels(Idents(combined))[i])
  this_scina_obj <- combined[, this_cluster_cells]
  Idents(this_scina_obj) <- this_scina_obj$scina_cell_type
  # select the largest overlapped cell type name as annotation of this seurat cluster, the unknown is ignored
  if ("unknown" %in% levels(Idents(this_scina_obj))) {
    scina_obj_ident_length <- length(levels(Idents(this_scina_obj))) - 1
    max_ct <- names(which.max(table(Idents(this_scina_obj))[1:scina_obj_ident_length]))
  } else {
    scina_obj_ident_length <- length(levels(Idents(this_scina_obj)))
    max_ct <- names(which.max(table(Idents(this_scina_obj))[1:scina_obj_ident_length]))
  }
  annotate_cell_type_name <- c(annotate_cell_type_name, max_ct)
}


levels(annotate_cell_type) <- annotate_cell_type_name
combined <- AddMetaData(combined, as.factor(annotate_cell_type), col.name = 'scina_cell_type2')
Idents(combined) <- combined$scina_cell_type
p2 <- Plot.cluster2D(combined,reduction.method = "umap",pt_size = 0.5, txt = "SCINA cell type")
p2

png(
  paste("SCINA_raw_cell_type.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
p2
dev.off()


Idents(combined) <- combined$scina_cell_type2
p3 <- Plot.cluster2D(combined,reduction.method = "umap",pt_size = 0.5, txt = "SCINA cell type")


png(
  paste("SCINA_cell_type.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
p3
dev.off()
p3

# Save marker gene plots, only save used
# Note: created a tmp object for upper case rownames
i <- 1
total_scina_cell_type <- levels(as.factor(combined$scina_cell_type2))
tmp_combined <- combined
DefaultAssay(tmp_combined) <- "RNA"
rownames(tmp_combined@assays$RNA@data) <- toupper(rownames(tmp_combined@assays$RNA@data))
for (i in 1:length(total_scina_cell_type)) {
  this_ct <- total_scina_cell_type[i]
  this_markers <- signatures[[this_ct]]
  for (j in this_markers) {
    if (j %in% toupper(rownames(tmp_combined))) {
      p_scina <- FeaturePlot(tmp_combined, j, ncol = 1)
      png(
        paste(
          "SCINA_",
          this_ct,
          "_marker_",
          j,
          ".png",
          sep = ""
        ),
        width = 1500,
        height = 1500,
        res = 300
      )
      print(p_scina)
      dev.off()
    }
  }
}


```
## SCINA cell type UMAP 

```{r,echo=F,eval=T,message=F,warning=F,error=F}

Idents(combined) <- combined$scina_cell_type2
p1 <- Plot.cluster2D(combined,reduction.method = "umap",pt_size = 0.5, txt = "SCINA cell type")
p1

png(
  paste("SCINA_cell_type.png", sep = ""),
  width = 2500,
  height = 2000,
  res = 300
)
print(p1)
dev.off()
```
## SCINA cell type UMAP with label

```{r,echo=F,eval=T,message=F,warning=F,error=F}

p2 <- DimPlot(
  combined,
  reduction = "umap",
  cols = as.vector(palette36.colors(36)[-2]),
  label = T,
  label.size = 3,
  repel = T
)

p2 
png(
  paste("SCINA_cell_type_with_label.png", sep = ""),
  width = 2500,
  height = 2000,
  res = 300
)
print(p2)
dev.off()

```

## SCINA cell type UMAP with label box

```{r,echo=F,eval=T,message=F,warning=F,error=F}

p3 <- DimPlot(
  combined,
  reduction = "umap",
  cols = as.vector(palette36.colors(36)[-2]),
  label = T,
  label.size = 3,
  label.box = T,
  repel = T
)

p3 
png(
  paste("SCINA_cell_type_with_label_box.png", sep = ""),
  width = 3000,
  height = 2000,
  res = 300
)
print(p3)
dev.off()
```


## SCINA cell type UMAP plot split by condition

Number of cells for each cell type predicted from SCINA:
 
```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}
table(combined$Condition, combined$scina_cell_type2)

```


```{r,echo=F,eval=T,message=F,warning=F,error=F, fig.width=12,fig.height=8}

Idents(combined) <- combined$scina_cell_type2

p3 <-
  DimPlot(
    combined,
    reduction = "umap",
    split.by = "Condition",
    cols = as.vector(palette36.colors(36)[-2]),
    pt.size = 0.7,
    ncol = 2
  )


png(
  paste("SCINA_cell_type_split_by_condition.png", sep = ""),
  width = 4000,
  height = 2000,
  res = 300
)
p3
dev.off()
p3

```

```{r,echo=F,eval=T,message=F,warning=F,error=F}


```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

```


```{r,echo=F,eval=F,message=F,warning=F,error=F}

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}

```

```{r,echo=F,eval=F,message=F,warning=F,error=F}
# Manually save the rds, do not save on knitr
saveRDS(combined, "combined.rds")
```


