---
title: "4. Cell type annotation"
author: "Cankun Wang"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
library(cowplot)
library(ggplot2)
library(patchwork)
library(here)
library(qs)
library(Polychrome)
library(RColorBrewer)
library(tidyverse)
library(SCINA)
library(preprocessCore)
library(garnett)
```


# Set working dir

```{r,echo=F,eval=T,message=F,warning=F,error=F}
here::i_am("4_cell_type_annotation.Rmd")
here::set_here()
print(paste("Current working directory:", here::here()))

```

# Load data

```{r,echo=F,eval=T,message=F,warning=F,error=F}
source("functions.R")

combined <- qs::qread('combined.qsave')

DefaultAssay(combined) <- "RNA"
Idents(combined) <- combined$seurat_clusters

custom_color <-
  as.character(palette36.colors(36)[-2])
two_color <- c('#C0C0C0', '#B00D23')

provided_marker <- read_csv("../provided_marker.csv") %>%
  mutate(cell_type = factor(cell_type))

provided_marker
```

```{r,echo=F,eval=F,message=FALSE,warning=F}

```


```{r,echo=F,eval=T,message=T,warning=F,error=F}

Idents(combined) <- combined$orig.ident
DimPlot(
  combined,
  reduction = "umap",
  cols = custom_color,
  label = T,
  pt.size = 0.4,
  repel = T,
  label.box = T
)

```

# Annotate by each sample

Criteria:

- Having all marker gene expression > 0.1
- If a cell matches multiple cell type based on marker expression, use the cell type label with the highest expression


We used Garnett to 

# Annotate by sample

We used the custom markers in the /marker folder to perform cell type annotation by each sample. Here we used Garnett (Pliner et al, 2019) tool and wrapped in a custom function 'generate_cell_type' (in functions.R). The parameters were adjusted to allow more unstrict annotation due to our in vitro experiment design. Once the automated process complete, we mannually examed and evaluated the annotation results.

Not showing in this markdown file, we have also tested some other tools like scCATCH, singleR, SCINA. All these other tools would generate a lot of misannotation and they can not handle negative marker genes.

```{r,echo=F,eval=T,message=F,warning=F,error=F}


sample_df <-
  read.csv("../marker/garnett_sample_list.csv", header = T)
sample_df

i = 11

for (i in seq_len(nrow(sample_df))) {
  this_sample <- sample_df$sample[i]
  print(this_sample)
  this_day <- sample_df$day[i]
  marker_file_path <- paste0("../marker/", this_day, "_marker.txt")
  this_combined <- subset(combined, idents = this_sample)
  
  garnett_obj <-
    generate_cell_type(seurat_obj = this_combined, marker_file_path = marker_file_path)
  
  this_combined <-
    AddMetaData(this_combined, as.factor(garnett_obj$cell_type), col.name = "garnett_cell_type")
  this_combined <-
    AddMetaData(this_combined,
                as.factor(garnett_obj$cluster_ext_type),
                col.name = "garnett_cell_type_ext")
  Idents(this_combined) <- this_combined$garnett_cell_type
  
  p1 <-
    DimPlot(
      this_combined,
      reduction = 'umap',
      pt.size = 0.7,
      cols = custom_color
    )
  
  Idents(this_combined) <- this_combined$garnett_cell_type_ext
  
  p2 <-
    DimPlot(
      this_combined,
      reduction = 'umap',
      pt.size = 0.7,
      cols = custom_color
    )
  
  qs::qsave(this_combined,
            paste0("../marker/", this_sample, "_garnett.qsave"))
  
  png(
    paste0("../marker/", this_sample, "_p1.png"),
    width = 3000,
    height = 2000,
    res = 300
  )
  p1
  dev.off()
  
  png(
    paste0("../marker/", this_sample, "_p2.png"),
    width = 3000,
    height = 2000,
    res = 300
  )
  p2
  dev.off()
  
  
}




```




```{r,echo=F,eval=F,message=F,warning=F,error=F}
qs::qsave(combined,'combined.qsave')
```


```{r,echo=F,eval=T,message=F,warning=F,error=F}

```
